### 2015-11-02 paper/chaining.tex
Adding initial version of writeup.

%!TEX TS-program = pdflatexmk

%\documentclass{sig-alt-hotnetscr}
\documentclass[12pt]{article}

\input{setup}

\begin{document}
\clubpenalty=10000
\widowpenalty = 10000
\thispagestyle{empty}

\input{title}
\input{protocol}
\input{discussion}

\bibliographystyle{abbrv}

\bibliography{bibinfo2}

%\endgroup

\end{document}