### 2016-02-28 paper/related.tex
corrected two typos

\section{Related Work}\label{sec-related}
Privacy-preserving computation has been studied extensively. An overview
of the approaches taken by the cryptographic-research community is provided by
Perry {\it et al.}~\cite{pgfw-scn2014}; an overview of of the data-mining
approach is provided by Aggarwal and Yu~\cite{ay2008}.  Kamara~\cite{kamara}
and Kroll {\it et al.}~\cite{kroll} used cryptographic protocols to 
achieve privacy and accountability in the surveillance of known targets.
Segal {\it et al.}~\cite{sff-foci2014} formulated the openness principle that
we have followed and were the first to design privacy-preserving protocols for
the surveillance of unknown targets.  Kearns {\it et al.}~\cite{krwy-pnas16}
present efficient graph-search algorithms that distinguish targeted users from
untargeted users while preserving the privacy of direct contacts of untargeted
users; unlike our privacy-preserving contact-chaining algorithms, which rely
on cryptographic techniques, their search algorithms rely on 
differential-privacy techniques.