### 2016-02-29 paper/submission.tex

\documentclass[USenglish,oneside,twocolumn]{article}
\usepackage{etex}
\usepackage[utf8]{inputenc}%(only for the pdftex engine)
%\RequirePackage[no-math]{fontspec}%(only for the luatex or the xetex engine)
\usepackage[big]{dgruyter_NEW}

\DOI{foobar}

\cclogo{\includegraphics{by-nc-nd.pdf}}

\DeclareMathOperator{\Enc}{Enc}

\DeclareMathOperator{\Dec}{Dec}

\DeclareMathOperator{\de}{deg}

\usepackage{subcaption}
\usepackage{pgfplots}

\begin{document}

  %\author*[1]{Aaron Segal}
  \author[1]{Names withheld}

  %\author[2]{Bryan Ford}

  %\author[3]{Joan Feigenbaum}

  %\affil[1]{Yale Univ., E-mail: Aaron.Segal@yale.edu}
  \affil[1]{Affiliations and email addresses withheld}

  %\affil[2]{EPFL, E-mail: Bryan.Ford@epfl.ch}

  %\affil[3]{Yale Univ., E-mail: Joan.Feigenbaum@yale.edu}

  \title{\huge Open, privacy-preserving protocols for lawful surveillance}

  \runningtitle{Privacy-preserving, lawful surveillance}

  %\subtitle{...}

  \begin{abstract}
{As demonstrated, {\it e.g.}, by Snowden's revelation of controversial NSA 
activity and by Apple's refusal to obey a Justice Department's order to break 
an iPhone security mechanism, the question of how government agencies can
acquire actionable, useful information about legitimate but unknown targets 
without intruding upon the electronic activity of innocent parties is 
extremely important.  We address this question by providing experimental
evidence that actionable, useful information can indeed be obtained in a manner
that preserves the privacy of innocent parties and that holds government
agencies accountable.  In particular, we present practical, privacy-preserving
protocols for two operations that law-enforcement and intelligence agencies
have used effectively: {\it set intersection} and {\it contact chaining}.}
  \end{abstract}
  \keywords{Privacy, Secure multiparty computation, Surveillance}
%  \classification[PACS]{}
 % \communicated{...}
 % \dedication{...}

  \journalname{Proceedings on Privacy Enhancing Technologies}
\DOI{Editor to enter DOI}
  \startpage{1}
  \received{February 29, 2016}
  %\revised{..}
  %\accepted{..}

  \journalyear{2016}
  \journalvolume{2016}
  \journalissue{4}

\maketitle

\input{intro}
\input{open}
\input{intersection}
\input{chaining}
\input{related}
\input{conc}
%\input{acks}

%% ###################################################################

\begin{thebibliography}{99}
%\bibitem{journal-1} A.~P.~Raposo, H.~J.~Weber, D.~E.~Alvarez--Castillo, M.~Kirchbach, Cent. Eur. J. Phys. 5, 253 (2007)
%\bibitem{journal-2} J.~Barth et al. (SAPHIR Collaboration), Phys. Lett. B 572, 127 (2003)
%\bibitem{journal-3} S.~Chekanov et al., Eur. Phys. J. C 51, 289 (2007)
%\bibitem{journal-4} K.~Malarz, Postepy Fizyki 57, 235 (2006) (in Polish)
%\bibitem{journal-5} G.~Meng, Cent. Eur. J. Phys., DOI:10.2478/s11534-007-0038-1
%\bibitem{journal-6} R.~Hegselmann, U.~Krause, Journal of Artificial Societies and Social Simulation (2006), https://urldefense.proofpoint.com/v2/url?u=http-3A__jasss.soc.surrey.ac.uk_9_3_10.html&d=AwIGaQ&c=-dg2m7zWuuDZ0MUcV7Sdqw&r=TO0xacJWWmvm7m0roFn_liAsIL2PkxELeLGL8JNJXFw&m=Ka2FvolwqKNW8vu0FjJrZZz-0NBFnF_Zu-dQp-fjTto&s=370GY1pBTDyr9NSQToGqkNR8K-Xy1z0bsX64GpcLiPk&e= 
%\bibitem{journal-7} A.~Dybala, Cent. Eur. J. Chem. (in press)
%\bibitem{journal-8} A.~Dybala, Przeglad chemiczny (in Polish, in press)
%\bibitem{book} M.~Lister, Fundamentals of Operating Systems, 3rd edition (Springer-Verlag, New York, 1984)
%\bibitem{chapter} C.~K.~Clenshaw, K.~Lord, In: B.~K.~P.~Scaife (Ed.), Studies in Numerical Analysis (Academic Press, London and New York, 1974) 95
%\bibitem{arxiv-1} M.~Majewski, K.~Malarz, arXiv:cond-mat/0609635v2 [cond-mat.stat-mech]
%\bibitem{arxiv-2} J.~A.~C.~E.~Solano, arXiv:0707.1343v1 [astro-ph]
\bibitem{sff-foci2014} A.~Segal, B.~Ford, J.~Feigenbaum, 
In: J. Crandall and V. Paxson (Eds.), $4^{th}$ Workshop on Free and 
Open Communications on the Internet, August 18, 2014, San Diego, 
USA (USENIX, Berkeley, 2014)

\bibitem{anderson13cell} N.~Anderson, arstechnica, August 29, 2013.

\bibitem{bieber-penn2002} M.~Bieber, Univ.~Penn.~Law Review, 150, 1079 (2002)

\bibitem{dissent} H.~Corrigan-Gibbs, B.~Ford, In: A.~D.~Keromytis and 
V.~Shmatikov (Eds.), $17^{th}$ Conference on Computer and Communications 
Security, October 4-8, 2010, Chicago, USA (ACM, New York, 2010)

\bibitem{verdict} H.~Corrigan-Gibbs, D.~I.~Wolinsky, B.~Ford, In: 
S.~King (Ed.), $22^{nd}$ Security Symposium, August 14-16, 2013, 
Washington, D.C., USA (USENIX, Berkeley, 2013)

\bibitem{snapnets} J.~Leskovec, A.~Krevl, SNAP Datasets: 
Stanford Large Network Dataset Collection, 
http://snap.stanford.edu/data

\bibitem{mobiforge} MobiForge, Global mobile statistics 2014 Part A: 
Mobile subscribers; handset market share; mobile operators,\\ 
https://mobiforge.com/research-analysis/global-mobile-\\
statistics-2014-part-a-mobile-subscribers-handset-market-\\
share-mobile-operators\#topmnos

\bibitem{pgfw-scn2014} J. Perry, D. Gupta, J. Feigenbaum, R. Wright,
In: M. Abdalla and R. De Prisco (eds.), $9^{th}$ Conference on Security and 
Cryptography for Networks, September 3-5, 2014, Amalfi, Italy (Springer,
Berlin, 2014)

\bibitem{ay2008} C.~Aggarwal, P.~Yu, In: C.~Aggarwal and P.~Yu (Eds.), 
Privacy-Preserving Data Mining: Models and Algorithms 
(Springer, Berlin, 2008) 11

\bibitem{kamara} S.~Kamara, In: M.~Brenner and M.~Smith (eds.), 
$2^{nd}$ Workshop on Applied Homomorphic Cryptography and Encrypted Computing, 
March 7, 2014, Barbados, West Indies (Springer, Berlin, 2014)

\bibitem{kroll} J.~Kroll, E. Felten, D. Boneh, Secure protocols for
accountable warrant execution, April 2, 2014,\\
https://freedom-to-tinker.com/blog/felten/\\
secure-protocols-for-accountable-warrant-execution/

\bibitem{krwy-pnas16} M. Kearns, A. Roth, Z. S. Wu, G. Yaroslavtsev, 
Proc.~Natl.~Acad.~Scis. 113, 913 (2016)

%\bibitem{proceedings} A.~Kaczanowski, K.~Malarz, K.~Kulakowski, In: T.~E.~Simos (Ed.), International Conference of Computational Methods in Science and Engineering, Sep. 12-16, 2003, Kastoria, Greece (World Scientific, Singapore 2003) 258
%\bibitem{thesis} A.~J.~Agutter, Ph.D. thesis, Edinburgh University (Edinburgh, UK, 1995)
%\bibitem{newspaper-1} A.~Sherwin, The Times, Jul. 13, 2007, 1
%\bibitem{newspaper-2} M.~Dzierzanowski, Wprost, Jul. 8, 2007, 18 (in Polish)
%\bibitem{patent} Philip Morris Inc., European patent application 0021165 A1, Jan. 7, 1981
%\bibitem{standard-1} ISO 2108:1992, Information and documentation --- International standard book numbering (ISBN)
%\bibitem{standard-2} ISO/TR 9544:1988, Information processing --- Computer-assisted publishing --- Vocabulary
\end{thebibliography}

\end{document}