### 2016-02-29 paper/submission.tex
Introduction edits

\documentclass[USenglish,oneside,twocolumn]{article}
\usepackage{etex}
\usepackage[utf8]{inputenc}%(only for the pdftex engine)
%\RequirePackage[no-math]{fontspec}%(only for the luatex or the xetex engine)
\usepackage[big]{dgruyter_NEW}

\DOI{foobar}

\cclogo{\includegraphics{by-nc-nd.pdf}}

\DeclareMathOperator{\Enc}{Enc}

\DeclareMathOperator{\Dec}{Dec}

\DeclareMathOperator{\de}{deg}

\usepackage{subcaption}
\usepackage{pgfplots}

\begin{document}

  %\author*[1]{Aaron Segal}
  \author[1]{Names withheld}

  %\author[2]{Bryan Ford}

  %\author[3]{Joan Feigenbaum}

  %\affil[1]{Yale Univ., E-mail: Aaron.Segal@yale.edu}
  \affil[1]{Affiliations and email addresses withheld}

  %\affil[2]{EPFL, E-mail: Bryan.Ford@epfl.ch}

  %\affil[3]{Yale Univ., E-mail: Joan.Feigenbaum@yale.edu}

  \title{\huge Open, privacy-preserving protocols for lawful surveillance}

  \runningtitle{Privacy-preserving, lawful surveillance}

  %\subtitle{...}

  \begin{abstract}
{The question of how government agencies can
acquire actionable, useful information about legitimate but unknown targets 
without intruding upon the electronic activity of innocent parties is 
extremely important.  We address this question by providing experimental
evidence that actionable, useful information can indeed be obtained in a manner
that preserves the privacy of innocent parties and that holds government
agencies accountable.  In particular, we present practical, privacy-preserving
protocols for two operations that law-enforcement and intelligence agencies
have used effectively: {\it set intersection} and {\it contact chaining}.
Experiments with our protocols suggest
that privacy-preserving contact chaining can perform
a 3-hop privacy-preserving graph traversal producing 27,000 ciphertexts
in under two minutes.
These ciphertexts are usable in turn via private set
intersection to pinpoint potential unknown targets within a body of 150,000
total ciphertexts within 10 minutes,
without exposing personal information about non-targets.
}
  \end{abstract}
  \keywords{Privacy, Secure multiparty computation, Surveillance}
%  \classification[PACS]{}
 % \communicated{...}
 % \dedication{...}

  \journalname{Proceedings on Privacy Enhancing Technologies}
\DOI{Editor to enter DOI}
  \startpage{1}
  \received{February 29, 2016}
  %\revised{..}
  %\accepted{..}

  \journalyear{2016}
  \journalvolume{2016}
  \journalissue{4}

\maketitle

\input{intro}
\input{open}
\input{intersection}
\input{chaining}
\input{related}
\input{conc}
%\input{acks}

%% ###################################################################

\begin{thebibliography}{99}
%\bibitem{journal-1} A.~P.~Raposo, H.~J.~Weber, D.~E.~Alvarez--Castillo, M.~Kirchbach, Cent. Eur. J. Phys. 5, 253 (2007)
%\bibitem{journal-2} J.~Barth et al. (SAPHIR Collaboration), Phys. Lett. B 572, 127 (2003)
%\bibitem{journal-3} S.~Chekanov et al., Eur. Phys. J. C 51, 289 (2007)
%\bibitem{journal-4} K.~Malarz, Postepy Fizyki 57, 235 (2006) (in Polish)
%\bibitem{journal-5} G.~Meng, Cent. Eur. J. Phys., DOI:10.2478/s11534-007-0038-1
%\bibitem{journal-6} R.~Hegselmann, U.~Krause, Journal of Artificial Societies and Social Simulation (2006), https://urldefense.proofpoint.com/v2/url?u=http-3A__jasss.soc.surrey.ac.uk_9_3_10.html&d=AwIGaQ&c=-dg2m7zWuuDZ0MUcV7Sdqw&r=TO0xacJWWmvm7m0roFn_liAsIL2PkxELeLGL8JNJXFw&m=Ka2FvolwqKNW8vu0FjJrZZz-0NBFnF_Zu-dQp-fjTto&s=370GY1pBTDyr9NSQToGqkNR8K-Xy1z0bsX64GpcLiPk&e= 
%\bibitem{journal-7} A.~Dybala, Cent. Eur. J. Chem. (in press)
%\bibitem{journal-8} A.~Dybala, Przeglad chemiczny (in Polish, in press)
%\bibitem{book} M.~Lister, Fundamentals of Operating Systems, 3rd edition (Springer-Verlag, New York, 1984)
%\bibitem{chapter} C.~K.~Clenshaw, K.~Lord, In: B.~K.~P.~Scaife (Ed.), Studies in Numerical Analysis (Academic Press, London and New York, 1974) 95
%\bibitem{arxiv-1} M.~Majewski, K.~Malarz, arXiv:cond-mat/0609635v2 [cond-mat.stat-mech]
%\bibitem{arxiv-2} J.~A.~C.~E.~Solano, arXiv:0707.1343v1 [astro-ph]
%\bibitem{sff-foci2014} A.~Segal, B.~Ford, J.~Feigenbaum, 
%In: J. Crandall and V. Paxson (Eds.), $4^{th}$ Workshop on Free and 
%Open Communications on the Internet, August 18, 2014, San Diego, 
%USA (USENIX, Berkeley, 2014)

\bibitem{sff-foci2014} Bandits, Catching Bandits and Only Bandits:
Privacy-Preserving Intersection Warrants for Lawful Surveillance,
\url{https://drive.google.com/file/d/0BwK1bcoczINtbUhPaFI0Mk1DVk0/view}
(2016)

\bibitem{anderson13cell} N.~Anderson, arstechnica, August 29, 2013.

\bibitem{bieber-penn2002} M.~Bieber, Univ.~Penn.~Law Review, 150, 1079 (2002)

\bibitem{dissent} H.~Corrigan-Gibbs, B.~Ford, In: A.~D.~Keromytis and 
V.~Shmatikov (Eds.), $17^{th}$ Conference on Computer and Communications 
Security, October 4-8, 2010, Chicago, USA (ACM, New York, 2010)

\bibitem{verdict} H.~Corrigan-Gibbs, D.~I.~Wolinsky, B.~Ford, In: 
S.~King (Ed.), $22^{nd}$ Security Symposium, August 14-16, 2013, 
Washington, D.C., USA (USENIX, Berkeley, 2013)

\bibitem{snapnets} J.~Leskovec, A.~Krevl, SNAP Datasets: 
Stanford Large Network Dataset Collection, 
http://snap.stanford.edu/data

\bibitem{mobiforge} MobiForge, Global mobile statistics 2014 Part A: 
Mobile subscribers; handset market share; mobile operators,\\ 
https://mobiforge.com/research-analysis/global-mobile-\\
statistics-2014-part-a-mobile-subscribers-handset-market-\\
share-mobile-operators\#topmnos

\bibitem{pgfw-scn2014} J. Perry, D. Gupta, J. Feigenbaum, R. Wright,
In: M. Abdalla and R. De Prisco (eds.), $9^{th}$ Conference on Security and 
Cryptography for Networks, September 3-5, 2014, Amalfi, Italy (Springer,
Berlin, 2014)

\bibitem{ay2008} C.~Aggarwal, P.~Yu, In: C.~Aggarwal and P.~Yu (Eds.), 
Privacy-Preserving Data Mining: Models and Algorithms 
(Springer, Berlin, 2008) 11

\bibitem{kamara} S.~Kamara, In: M.~Brenner and M.~Smith (eds.), 
$2^{nd}$ Workshop on Applied Homomorphic Cryptography and Encrypted Computing, 
March 7, 2014, Barbados, West Indies (Springer, Berlin, 2014)

\bibitem{kroll} J.~Kroll, E. Felten, D. Boneh, Secure protocols for
accountable warrant execution, April 2, 2014,\\
https://freedom-to-tinker.com/blog/felten/\\
secure-protocols-for-accountable-warrant-execution/

\bibitem{krwy-pnas16} M. Kearns, A. Roth, Z. S. Wu, G. Yaroslavtsev, 
Proc.~Natl.~Acad.~Scis. 113, 913 (2016)

\bibitem{elgamal} T.~ElGamal, IEEE Transactions on Information Theory, IT-31 No. 4 (1985)

\bibitem{pohlighellman} S.~Pohlig, M.~Hellman, IEEE Transactions on Information Theory, IT-24 No. 1 (1978)

\bibitem{techdirt} T.~Cushing, techdirt.com, Jul. 3rd, 2014, \url{https://www.techdirt.com/articles/20140629/16130227727/nsa-appears-to-be-chaining-calls-using-phone-numbers-one-hop-out-as-new-originating-selectors.shtml}

%\bibitem{proceedings} A.~Kaczanowski, K.~Malarz, K.~Kulakowski, In: T.~E.~Simos (Ed.), International Conference of Computational Methods in Science and Engineering, Sep. 12-16, 2003, Kastoria, Greece (World Scientific, Singapore 2003) 258
%\bibitem{thesis} A.~J.~Agutter, Ph.D. thesis, Edinburgh University (Edinburgh, UK, 1995)
%\bibitem{newspaper-1} A.~Sherwin, The Times, Jul. 13, 2007, 1
%\bibitem{newspaper-2} M.~Dzierzanowski, Wprost, Jul. 8, 2007, 18 (in Polish)
%\bibitem{patent} Philip Morris Inc., European patent application 0021165 A1, Jan. 7, 1981
%\bibitem{standard-1} ISO 2108:1992, Information and documentation --- International standard book numbering (ISBN)
%\bibitem{standard-2} ISO/TR 9544:1988, Information processing --- Computer-assisted publishing --- Vocabulary
\end{thebibliography}

\end{document}