### 2016-07-12 paper/submission.tex
Put citations from bibitems into chaining.bib

%\documentclass[USenglish,oneside,twocolumn]{article}
\documentclass{sig-alternate-05-2015}

\usepackage{etex}
\usepackage[utf8]{inputenc}%(only for the pdftex engine)
%\RequirePackage[no-math]{fontspec}%(only for the luatex or the xetex engine)
%\usepackage[big]{dgruyter_NEW}
\usepackage{amsmath}
\usepackage{hyperref}

%\DOI{foobar}

%\cclogo{\includegraphics{by-nc-nd.pdf}}

\DeclareMathOperator{\Enc}{Enc}

\DeclareMathOperator{\Dec}{Dec}

\DeclareMathOperator{\de}{deg}

\usepackage{subcaption}
\usepackage{pgfplots}

  \title{Open, privacy-preserving protocols for lawful surveillance}
  \author{
	Aaron Segal and Joan Feigenbaum \\
	Yale University \\
	New Haven, CT, USA\\
	\{aaron.segal,joan.feigenbaum\}@yale.edu\\
	\and
	Bryan Ford \\
	Swiss Federal Institute of Technology (EPFL) \\
	Lausanne, Switzerland\\
	bryan.ford@epfl.ch}
  \date{}

  %\author*[1]{Aaron Segal}
  %\author[1]{Names withheld}

  %\author[2]{Bryan Ford}

  %\author[3]{Joan Feigenbaum}

  %\affil[1]{Yale Univ., E-mail: Aaron.Segal@yale.edu}
  %\affil[1]{Affiliations and email addresses withheld}

  %\affil[2]{EPFL, E-mail: Bryan.Ford@epfl.ch}

  %\affil[3]{Yale Univ., E-mail: Joan.Feigenbaum@yale.edu}

  %\title{\huge Open, privacy-preserving protocols for lawful surveillance}

  %\runningtitle{Privacy-preserving, lawful surveillance}

  %\subtitle{...}

\begin{document}

\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
%\doi{10.475/123_4}

% ISBN
%\isbn{123-4567-24-567/08/06}

%Conference
\conferenceinfo{WPES '16}{October 24, 2016, Vienna, Austria}

%\acmPrice{\$15.00}

\maketitle

  \begin{abstract}
{The question of how government agencies can
acquire actionable, useful information about legitimate but unknown targets 
without intruding upon the electronic activity of innocent parties is 
extremely important.  We address this question by providing experimental
evidence that actionable, useful information can indeed be obtained in a manner
that preserves the privacy of innocent parties and that holds government
agencies accountable.  In particular, we present practical, privacy-preserving
protocols for two operations that law-enforcement and intelligence agencies
have used effectively: {\it set intersection} and {\it contact chaining}.
Experiments with our protocols suggest
that privacy-preserving contact chaining can perform
a 3-hop privacy-preserving graph traversal producing 27,000 ciphertexts
in under two minutes.
These ciphertexts are usable in turn via privacy-preserving set
intersection to pinpoint potential unknown targets within a body of 150,000
total ciphertexts within 10 minutes,
without exposing personal information about non-targets.
}
  \end{abstract}
%  \keywords{Privacy, Secure multiparty computation, Surveillance}
%  \classification[PACS]{}
 % \communicated{...}
 % \dedication{...}

  %\journalname{Proceedings on Privacy Enhancing Technologies}
  %\DOI{Editor to enter DOI}
  %\startpage{1}
  %\received{February 29, 2016}
  %\revised{..}
  %\accepted{..}

  %\journalyear{2016}
  %\journalvolume{2016}
  %\journalissue{4}

\maketitle

\input{intro}
\input{open}
\input{intersection}
\input{chaining}
\input{related}
\input{conc}
%\input{acks}

%% ###################################################################
\bibliography{chaining}
\bibliographystyle{plain}

\end{document}